Human capital mediates natural selection in contemporary humans

### 2021-12-01 why-natural-selection.Rmd
Focus on motivating the theory as a reason for sample-splitting. Reorder intro to put theory first.

This
study uses data from UK Biobank [@bycroft2018uk] to learn more about contemporary
natural selection. We test for
natural selection on 33 different polygenic scores by estimating their
correlation with fertility. We extend the analysis over two generations, using
data on respondents' number of siblings as well as their number of children.
This is interesting because consistent natural selection over multiple
generations could lead to substantive effects in the long run. We then examine
reproductive rates in different subgroups of the population. 
We find selection effects on many polygenic scores. Effects are largely consistent
across generations.

Next, we examine correlations with fertility in different subgroups. Across
the board, selection effects are stronger in groups with lower
income and less education, among younger parents, people not living with a partner,
and people with more lifetime sexual partners. Outside these groups, effects are
weaker and often statistically insignificant. In some subgroups, the
direction of selection is even reversed.

We then show that a parsimonious model of human capital, education and fertility
choices can give rise to these empirical results. To test the theory, we show
that polygenic scores' correlation with fertility is associated with their
correlation with education and earnings. Furthermore, controlling for education
and earnings reduces the correlation with fertility, suggesting that human
capital mediates natural selection. Thus, contemporary natural selection on
polygenic scores can be explained by their correlation with earnings-increasing
human capital.

>>>>

This
study uses data from UK Biobank [@bycroft2018uk] to learn more about contemporary
natural selection. We test for
natural selection on 33 different polygenic scores by estimating their
correlation with fertility. We extend the analysis over two generations, using
data on respondents' number of siblings as well as their number of children.
This is interesting because consistent natural selection over multiple
generations could lead to substantive effects in the long run. We then examine
reproductive rates in different subgroups of the population.

We find selection effects on many polygenic scores. Effects are largely consistent
across generations. The strength of natural selection on a polygenic score is
associated with that score's correlation with education and earnings: scores
that predict lower education and earnings are being selected for. Also, across
the board, polygenic scores have stronger relationships with fertility among
specific subgroups. Selection effects are stronger in groups with lower
income and less education, among younger parents, people not living with a partner,
and people with more lifetime sexual partners. Outside these groups, effects are
weaker and often statistically insignificant. In some subgroups, the
direction of selection is even reversed: polygenic scores predicting
higher education and earnings are associated with *higher* fertility.

We then show that a parsimonious model of human capital, education and
fertility choices can give rise to these empirical results. Thus, contemporary
natural selection on polygenic scores can be explained by their correlation with
earnings-increasing human capital.
